An SMT-based Approach to Coverability Analysis

### 2014-02-04 paper/workspace/main.tex
Split in sections, added references.

@@

\section{Preliminaries}

\begin{verbatim}
* Petri net:
 N = (S, T, E, M0), where
 S : places
 T : transitions
 E : edges
 M0: initial marking

* Property:
 P, linear arithmetic formula
\end{verbatim}
\iffalse
\begin{verbatim}
 Examples:
  x + y < 0
  x + y < 0 \\and x + z > 0
  x + y < 0 \\or  x + z > 0

 ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
\end{verbatim}
\fi

@@

\begin{verbatim}
* Code

Process 1:

    bit1 := false
    while true do
p1:   bit1 := true
p2:   while bit2 do skip od
p3:   <critical section>
      bit1 := false
    od

Process 2:

    bit2 := false
    while true do
q1:   bit2 := true
q2:   if bit1 then
q3:     bit2 := false
q4:     while bit1 do skip od
        goto q1
      fi
q5:   <critical section>
      bit2 := false
od

* Property: Process 1 and Process 2 are never in their respective critical section at the same time.

* Petri net:

* Subnet of Petri net:

\begin{center}
\begin{tikzpicture}[node distance=1cm]
  \node[transition, label=left:t1] (t1) {};
  \node[place, tokens=1, label=left:a, below of=t1] (a) {};
  \node[transition, label=left:t2, below of=a] (t2) {};
  \node[place, tokens=0, label=left:b, below of=t2] (b) {};
  \node[place, tokens=0, label=right:$\neg bit2$, right of=t2] (nbit2) {};
  \draw (t1) edge[pre, bend left, min distance=2cm, out=225, in=-45] (b);
  \draw (t1) edge[post] (a);
  \draw (t2) edge[pre] (nbit2);
  \draw (t2) edge[post] (nbit2);
  \draw (t2) edge[pre] (a);
  \draw (t2) edge[post] (b);
\end{tikzpicture}
\end{center}

\begin{verbatim}
* Property P: a + b <= 1
\end{verbatim}

@@

\begin{verbatim}
* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
\end{verbatim}

\begin{verbatim}
* Example:
  - State constraints C:

    a + b >= 2

@@

  \end{verbatim}
\begin{align*}
  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
  \right) \right) \land
    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
    \left( \bigwedge_{t \in T} t \ge 0 \right)
\end{align*}

@@

\newpage
\begin{verbatim}
* Example
  - State constraints C same as in the example for method Safety.

@@

  "At least one place in S is always marked"

* Subprocedure TrapConditions:

\newpage
\section{Method Invariant}

@@

\newpage
\begin{verbatim}
* Example:

\newpage
\begin{verbatim}
* Derivation of invariant I for Petri net N that satisfies property P:

@@

\newpage
\begin{verbatim}
* Subprocedure \mathcal{C'}:

\newpage
\section{Method Invariant with Minimization}

@@

\newpage
\begin{verbatim}
* Example:
  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant

@@

    a + b <= 1

* Subprocedures \mathcal{C'} and Inv are the same as
  in Method Invariant

@@

@@

    a     + nbit2 >= 1
    a + b + nbit2 <= 1

* Subprocedures TrapConditions and \Delta are the same as
  in Method Safety by Refinement

* Subprocedures \mathcal{C'} and Inv are the same as
  in Method Invariant

\iffalse

\newpage
\section{Benchmarks of \bfc}

We experimented on the benchmarks \ttt{cprover-PN} and
\ttt{cprover\_software\_analysis} available at \\
\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
Table~\ref{bfc-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
    Safety by Refinement                        & 23 &  0 & 54 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

\section{Benchmarks of \mist}

We experimented on the benchmarks available at \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
and \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
Table~\ref{mist-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
    Safety by Refinement                        & 20 &  0 &  7 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

\section{Benchmarks of \iic}

We experimented on the benchmarks available at \\
\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
Table~\ref{iic-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & pos & neg & don't know & TO \\
    \hline
    Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
    Safety by Refinement                        & 2 & 0 & 36 & 0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \iic}
\label{iic-experiments}
\end{table}

We constructed the following two invariants by applying the method
Invariant with Minimization.

\begin{itemize}
\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
\end{itemize}

\iic\ constructs the following two invariants.

\begin{itemize}
\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
\end{itemize}

\fi

>>>>

\nocite{DOsualdoSAS13}

@@

\input{sec_illustration}
\input{sec_preliminaries}

@@

@@

\begin{verbatim}
* Property of state constraints C: If C U {\neg P} is unsat then N |= P.

@@

\begin{align*}
  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
  \right) \right) \land
    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
    \left( \bigwedge_{t \in T} t \ge 0 \right)
\end{align*}

\begin{verbatim}
* Example

  - Code:

  Process 1:                     |      Process 2:
                                 |
      bit1 := false              |      bit2 := false
      while true do              |      while true do
  p1:   bit1 := true             |  q1:   bit2 := true
  p2:   while bit2 do skip od    |  q2:   if bit1 then
  p3:   <critical section>       |  q3:     bit2 := true
        bit1 := false            |  q4:     while bit1 do skip od
      od                         |          goto q1
                                 |        fi
                                 |  q5:   <critical section>
                                 |        bit2 := false
                                 |      od

  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.

  - Petri net:

\begin{verbatim}
  - State constraints C:

    a + b >= 2
\end{verbatim}

@@

\begin{verbatim}
* Subprocedure TrapConditions:

\begin{verbatim}
* Example

  - State constraints C same as in the example for method Safety.

@@

  "At least one place in S is always marked"
\end{verbatim}

\section{Method Invariant}

@@

\begin{verbatim}
* Subprocedure \mathcal{C'}:

\begin{verbatim}
* Derivation of invariant I for Petri net N that satisfies property P:

@@

\section{Method Invariant with Minimization}

@@

\begin{verbatim}
* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant

  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant

@@

    a + b <= 1
\end{verbatim}

@@

\begin{verbatim}
* Subprocedures TrapConditions and \Delta are the same as
  in Method Safety by Refinement

* Subprocedures \mathcal{C'} and Inv are the same as
  in Method Invariant
\end{verbatim}

@@

    a     + nbit2 >= 1
    a + b + nbit2 <= 1
\end{verbatim}

\input{sec_experiments}
\bibliographystyle{abbrv}
\bibliography{references}
